---
title: "Stats 100A HW 3 Milestone 3"
author: Mingye Wang, Kevin Liu
output: html_document
date: "2023-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.csv("Spring23Hwk3RawDatamergedcleanedFINAL-NO-NAMES.csv")
```

## Section 1

### 1.1 Data Introduction/Summary

Continuous variables: *walk*, *landscape*

Discrete with finite values: *pressure*, *transfer*, *stem*

Discrete with potentially infinitely many values: *hamburger*

```{r}
data_summary <- data.frame(matrix(ncol = 6, nrow = 5))
colnames(data_summary) <- colnames(data)[1:6]
rownames(data_summary) = c("min", "max", "mean", "std", "median")

for (b in colnames(data_summary))
  {
    data_summary[,b] <- c(min(data[,b]), max(data[,b]), mean(data[,b]), sd(data[,b]), median(data[,b]))
  }

data_summary
```

### 1.2 Empirical PMF for *pressure*

```{r}
pressure = data[,"pressure"]
pressure_pmf = cbind(c(1,2,3,4), c(mean(pressure == 1), mean(pressure == 2), mean(pressure == 3), mean(pressure == 4)))
colnames(pressure_pmf) <- c("X", "P(X)")

pressure_mean <- 0
for (i in seq(4))
{
  pressure_mean <- pressure_mean + pressure_pmf[i,1] * pressure_pmf[i,2]
}
pressure_mean

pressure_sd <- 0
for (i in seq(4))
{
  pressure_sd <- pressure_sd + pressure_pmf[i,1]^2 * pressure_pmf[i,2]
}
pressure_sd <- sqrt(pressure_sd - pressure_mean^2)
pressure_sd
```

### 1.3 Empirical PMF for *stem*

```{r}
stem = data[,"stem"]
stem_pmf = cbind(c(0,1), c(mean(stem == 0), mean(stem == 1)))
colnames(stem_pmf) <- c("X", "P(X)")

stem_mean <- 0
for (i in seq(2))
{
  stem_mean <- stem_mean + stem_pmf[i,1] * stem_pmf[i,2]
}
stem_mean

stem_sd <- 0
for (i in seq(2))
{
  stem_sd <- stem_sd + stem_pmf[i,1]^2 * stem_pmf[i,2]
}
stem_sd <- sqrt(stem_sd - stem_mean^2)
stem_sd
```

*stem* is a Bernoulli random variable, since it has only two outcomes: $0$ and $1$.

### 1.4a Joint PMF of *pressure* and *stem*

```{r}
pressure_stem <- cbind(pressure, stem)
ps_pmf <- matrix(ncol = 2, nrow = 4)
colnames(ps_pmf) <- c("S = 0", "S = 1")
rownames(ps_pmf) <- c("P = 1", "P = 2", "P = 3", "P = 4")
for (i in seq(4))
{
  for (j in seq(2))
  {
    ps_pmf[i,j] <- mean((pressure == i) & (stem == (j-1)))
  }
}
ps_pmf
```

### 1.4b Conditional empirical PMFs

```{r}
## Assume stem = 1
p_s1 <- ps_pmf[,2] / sum(ps_pmf[,2])
p_s1

mean_p_s1 <- 0
for (i in seq(4))
{
  mean_p_s1 <- mean_p_s1 + i * p_s1[i]
}


sd_p_s1 <- 0
for (i in seq(4))
{
  sd_p_s1 <- sd_p_s1 + (i^2) * p_s1[i]
}
sd_p_s1 <- sqrt(sd_p_s1 - mean_p_s1^2)
print(paste("Mean: ", mean_p_s1))
print(paste("Sd: ", sd_p_s1))

## Assume stem = 0
p_s0 <- ps_pmf[,1] / sum(ps_pmf[,1])
p_s0

mean_p_s0 <- 0
for (i in seq(4))
{
  mean_p_s0 <- mean_p_s0 + i * p_s0[i]
}

sd_p_s0 <- 0
for (i in seq(4))
{
  sd_p_s0 <- sd_p_s0 + (i^2) * p_s0[i]
}
sd_p_s0 <- sqrt(sd_p_s0 - mean_p_s0^2)
print(paste("Mean: ", mean_p_s0))
print(paste("Sd: ", sd_p_s0))

### Plots

plot(seq(4), p_s1, pch = 18, col = "green", main = "Figure 1: PMF of pressure given stem", xlab = "pressure", ylab = "P(pressure)")
points(seq(4), p_s0, pch = 19, col = "red")
abline(v=mean_p_s1, col="green")
abline(v=mean_p_s0, col="red")
legend("topleft",legend = c("stem = 1", "stem = 0"), pch = c(18, 19), col = c("green", "red"))
```

There is some indication that *pressure* and *stem* have some correlation, because we would expect the conditional pmf's to be the same if they were independent. Evidently this is not the case, as the pmf's have differing values for each value of *pressure* depending on the given value of *stem*.

## Section 2 

### 2.1 Empirical pmf of *hamburger*

```{r}
hamburger <- data[,"hamburger"]
burger_pmf <- matrix(ncol = 3, nrow = 9)
rownames(burger_pmf) <- c("0","1","2","3","4","5","6",">6", "total")
colnames(burger_pmf) <- c("obs", "freq", "th. prob.")

burger_pmf[,1] <- c(sum(hamburger == 0), sum(hamburger == 1), sum(hamburger == 2), sum(hamburger == 3), sum(hamburger == 4), sum(hamburger == 5), sum(hamburger == 6), sum(hamburger > 6), length(hamburger))

burger_pmf[,2] <- c(mean(hamburger == 0), mean(hamburger == 1), mean(hamburger == 2), mean(hamburger == 3), mean(hamburger == 4), mean(hamburger == 5), mean(hamburger == 6), mean(hamburger > 6), 1)

hb_mean <- sum(hamburger) / length(hamburger)
for (i in seq(7))
{
  burger_pmf[i,3] <- hb_mean ^ (i - 1) * exp(-1 * hb_mean) / factorial(i - 1)
}
burger_pmf[8, 3] <- 1 - sum(burger_pmf[1:7, 3])
burger_pmf[9, 3] <- 1

burger_pmf
```

### 2.2 Goodness of fit

```{r}
chi_sq <- matrix(ncol = 2, nrow = 9)
colnames(chi_sq) <- c("(O - E)^2", "(O - E)^2 / E")
for (i in seq(8))
{
  chi_sq[i,1] <- sqrt(length(hamburger) * burger_pmf[i, 3] - burger_pmf[i, 2])
  chi_sq[i,2] <- chi_sq[i,1] / (length(hamburger) * burger_pmf[i, 3])
}
burger_pmf <- cbind(burger_pmf, chi_sq)

burger_pmf
```



## Section 3

## Section 4



